Linear instability underlying synchronous waving of marine grass in flow

### 2014-09-21 Paper/grass.tex
changed the location of figures so that they appear in relevant para

\documentclass[aps,prl,twocolumn,superscriptaddress,10pt]{revtex4-1}  % for review and submission
%\documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
%\documentclass[letterpaper,10pt,prl,twocolumn,aps]{revtex4-1}
%\documentclass[aps,preprint,floatfix,prl]{revtex4-1}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{color}
\usepackage{longtable}
\usepackage{array}
\usepackage{amssymb}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

@@

\begin{abstract}
We show that the threshold flow condition for the spontaneous waving of marine grass arises due to the drag exerted by the vegetation on the flow. 
We show this by developing a self-consistent model of the flow past rigid vegetation. 
Linear stability analysis of the steady flow resulting from this model naturally leads to the threshold condition, and explains its mechanistic origin. 
Asymptotic analysis of the model for dense grass reveals two unstable modes. 
One of these modes shares many common features with the Kelvin-Helmholtz instability, but is demonstrated to be different from it. 
The predictions of the model for threshold condition and the resulting dominant frequency of waving agree with laboratory observations. 
\end{abstract}
\maketitle
%\section{Introduction}
Sea grasses exhibit a rich set of dynamical behavior due to their collective interaction with both steady and oscillatory flows.  
The resulting changes in hydrodynamic conditions can influence number of environmental processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production \cite{Fonseca87,Grizzle96,Nepf99,Nepf2012}. 
The simplest response of these flexible grass canopies to steady currents is the formation of coherent large amplitude oscillations, known as \monami \cite{AckermanOkubo93}.  
In this letter we provide a hydrodynamic mechanism for the onset of these coherent oscillations.
Current explanations of \monami invoke the existence of a shear layer near the top of the canopy \cite{Ghisal02,Raupach96} and its instability through a mechanism similar to the Kelvin-Helmholtz instability. 
This is thought to lead to coherent eddies over the sea grass canopy which then drives large amplitude synchronous oscillations.

\begin{figure}
\includegraphics[scale=1]{Grass_Base_Nepf_shear}
\caption{Comparison of steady flow profile from the experiments by Ghisalberti and Nepf\cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 
plant height = 13.7$\pm 0.2$ cm and blade width of 0.64 cm)
 and it's approximation with $U_0=7.28$ cm/s and $\delta = 5.02$ cm with our model. The canopy extends upto $y=h_g$ in the water column of depth $2H$. 
The steady velocity profile can be decomposed into three regions, (i) A parabolic profile in the unvegetated region, (ii) a uniform profile deep within the grass bed, and (iii) a boundary layer of thickness $\delta$ near the canopy top. 
The dependence of the boundary layer thickness (estimated numerically as $|U/U_y|$ at $y=\hg$) on the vegetation density parameter $\Rey \Ndg$ is shown in the inset.}
\label{basicflow}
\end{figure}

While the shear layer model appears to predict the frequency of \monami consistent with experimental observations, several aspect of existing theory remain unsatisfactory. 
(i) While the free shear layer is established due to the vegetation drag, linear stability analysis\cite{Raupach96} does not account for this, rendering this explanation inconsistent. 
(ii) Classical free shear flow is known to be unstable for all Reynolds number\cite{drazin}, whereas observations in the lab\cite{Ghisal02} and the field\cite{Grizzle96} indicate the existence of a threshold 
flow speed below which \monami is not observed. 
(iii)The thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the grass tip, is in many cases comparable to the unvegetated layer thickness,
and therefore inconsistent with the assumptions of the free shear layer instability.
Here we present a self-consistent mathematical model that account correctly for these effects 
while also explaining lab experiments and field observations.
%Using this model, the synchronous waving may be understood to be arising from the linear hydrodynamic instability of a uniform flow over the grass bed, such that the principal role of the vegetation is to exert a drag on the flow. 
%A threshold flow speed arises naturally for the onset of the instability; this threshold condition and the resulting waving frequency compares well with experimental observations.
%The vegetation drag introduces a damping mechanism in the flow, in addition to the viscous damping, and overcoming this damping is the primary consideration that sets the threshold flow.
%We deduce this numerically and through an asymptotic scaling analysis in the limit of dense grass.
\newline
The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying \monami because it realizes the shear layer structure of the flow. 
The flexibility of the grass blades, on the other hand, serves merely aid in visualizing the flow structure resulting from the instability\cite{Nepf2012}. 
It has been demonstrated that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels\cite{Ghisal02}. 
Therefore, for simplicity, we consider the grass bed to form a rigid structure that exchanges momentum with the surrounding fluid but does not deform. 
Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated\cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, 
the flow instability is similar to the instability of a shear layer.

\begin{figure}
\begin{center}
\includegraphics[]{Critical_Re_vs_delta_noshear} \\
\vspace{-6mm} \hspace{-3mm}
\includegraphics[]{K_vs_shear_width_noshear}
\end{center}
\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density
 parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are 
also included in the top panel. Inset compares experimental observations 
of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The experimental data in the inset in obtained from publications by 
Ghisalberti and Nepf \cite{Ghisal02} and Vivoni \cite{Vivoni98}. In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.}
\label{Re_vs_delta}
\end{figure}
We use a mean field model for the coupling between the flow and the canopy. 
The force exerted by the grass is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$,  
where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area. In the canopy the equations of continuity \& momentum balance can be written as 
%in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} as
\begin{equation}
\quad \grad\cdot\bu = 0,\hspace{3mm} \rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad P+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
\label{navier-stokes}
\end{equation}
%where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area, 
where $\rho$ the fluid density, $\mathbf{u}$ the velocity, 
$P$ the pressure, $\mu$ the turbulent viscosity and $\mathbf{g}$ the acceleration due to gravity. 
The Reynolds number of the flow based on the scale of a grass blade is expected to be O($10^2$-$10^3$), and hence the drag force is dominated by pressure and modeled as 
$\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn + C_{T}\rho u_{T}^{2}d\bt$ \shreyas{citation needed} 
where $d$ is average width of grass blade $C_{N}$ and $C_{T}$ are the normal and tangential drag coefficients respectively; $u_{T}$, $u_{N}$ are velocity components along and
normal to grass while $\bt,\bn$ being unit vector along and normal to grass. 
We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. 
In the field, both $C_N$ and $N_g$ vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N$ and $N_g$ to be constants. 
Based on previous experiments\cite{Vivoni98}, we take $C_N \approx 1$.

To understand the onset of \monami, we first calculate the steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ as a solution of \eqref{navier-stokes}, and examine its linear stability. 
Such a steady state flow driven by constant pressure gradient satisfies 
\begin{equation}
 -{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0
\label{base_equ}
\end{equation}
where $S(y)=1$ in the vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< 2H$. 
In the limit of large grass density, shear stress exerted by the bottom surface is estimated to be negligible in comparison with the vegetation drag and hence the bottom surface ($y=0$) is approximated by zero shear.   
A comparison of laboratory mean profile with the steady flow profile is shown in figure \eqref{basicflow}.
Flow within grass has approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from the balance between drag and with pressure gradient. 
Flow outside has a simple parabolic velocity profile due to the balance between viscous forces and the pressure gradient. 
Continuity of flow speed and shear stresses at canopy top gives rise to a boundary layer near the tip of the grass.

This boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and we identify it as a free shear layer\cite{Ghisal02} in the previous explanation of \monami. 
In our model the balance between viscous forces and the drag force in the shear layer 
%$C_NdN_g\delta^2u/\nu \sim O(1)$ 
together with the continuity of shear stress across the grass tip 
%$u \sim \delta U_o/H$
 leads to a shear layer thickness $\delta=(\Rey\tilde{N_g})^{-1/3}H$.
% and velocity $u \sim \delta U_0$ in the shear layer. 
Where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, and $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity. 
This dependence of the boundary layer thickness on the grass density $N_g$, shown in Figure \ref{basicflow}, gives us a way to systematically control it and investigate its effect on the instability mechanism.
The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$; minor departures from this scaling can be seen in the inset of figure \ref{basicflow}.

Considering small perturbations $u, v, p$ about the base profile $U$ and $P$ given by the solution of \eqref{base_equ}. Substituting them in equation \eqref{navier-stokes} and expanding  to linear order in the perturbation variable we obtain.
\begin{equation}
\begin{split}
\rho(u_t+U u_x+vU_y) &= -p_x+ {\mu}\nabla^2u-2S\rho C_{N}dN_{g}Uu, \\
\rho(v_t+ Uv_x) &= -p_y+ {\mu}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0.
\end{split}
\end{equation}
These equations can be non-dimensionalized using half channel height $H$, velocity $U_0$, and the associated advection time $H/U_0$. 
We are left with three non-dimensional parameters in the model, ie. $\Rey$, $\Ndg$, and the submergence ratio of the grass $h_g/H$. 
We also use $\delta/H$ in lieu of $\Ndg$ to parametrize the grass density and help elucidate the mechanism of the instability. 
With these scaling, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of 
the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \hat\phi(y) \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation. 
The growth rate $\sigma$ for a given wave number $k=2\pi /\lambda$ can be shown to satisfy the eigenvalue problem.
% provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$ through the solution of
\begin{equation}
\begin{split}
\left(D^2 -k^{2} \right)^2\phi &= \Rey \left[ \left({\sigma}+ikU\right) \left(D^2-k^2\right) -ikU_{yy}\right]\phi \\
&+D\left(2R \Ndg S U D \phi\right),
\label{Orr-somerfield}
\end{split}
\end{equation}
where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$.

@@

\begin{table*}
\rowcolors{3}{tableShade}{white}  %% start alternating shades from 3rd row
\renewcommand{\arraystretch}{1.4}
 \begin{tabular}{l|c|c|c}
			& Kelvin-Helmholtz 				& Mode 1 		& Mode 2 \\ \hline
 Base velocity profile 	& $U(y) = U_0 \tanh(y/\delta)$			& \multicolumn{2}{c}{$-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$} \\
 Domain 		& $-\infty < y < \infty$			& \multicolumn{2}{c}{$-1<y<1$} \\
 Inflection point	& exists at $y=0$				& \multicolumn{2}{c}{$U''(y)$ discontinuous at $y=\hg$} \\
 Shear layer thickness	& $\delta$					& \multicolumn{2}{c}{$\delta \sim  H\left(\Rey \Ndg \right)^{-1/3}$} \\
 Linearized dynamics	& Inviscid Rayleigh's equation			& \multicolumn{2}{c}{Equation \eqref{Orr-somerfield}} \\
 Critical parameters	& none						& $\Rey \propto \Ndg^{1/2}$ 	& $\Rey \propto \Ndg$ \\
 Most unstable $k$ as $\delta \to 0$	& $\propto \delta^{-1}$		& $\propto \delta^{-1}$	& $O(1)$ \\
 Mode localized?	& yes, near $y=0$				& ~~~~yes, near $y=\hg$~~~~			& no
 \end{tabular}
 \caption{Comparison between Kelvin-Helmholtz instability and the two modes resulting from solution of \ref{Orr-somerfield}.}
 \label{tab:comparison}
\end{table*}

To better understand the mechanism of waving and the origin of the threshold, we consider the behavior of the instability as a function of the vegetation drag.
The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical grass density discontinuously jumps and remains $O(1)$ (see figure \ref{Re_vs_delta}). 
To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in figure \ref{K_Re_sigma_set3}. 
The lowest $\Rey$ on the neutral curve sets the threshold. 
We observe that as $\Ndg$ increases, the unstable region splits into two. We refer to the unstable region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''. 
The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the grass density increases, causing the discontinuity in the most unstable mode.

>>>>

\documentclass[aps,prl,twocolumn,superscriptaddress,10pt]{revtex4-1}  % for review and submission
%\documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
%\documentclass[letterpaper,10pt,prl,twocolumn,aps]{revtex4-1}
%\documentclass[aps,preprint,floatfix,prl]{revtex4-1}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{color}
\usepackage{longtable}
\usepackage{array}
\usepackage{amssymb}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

@@

\begin{abstract}
We show that the threshold flow condition for the spontaneous waving of marine grass arises due to the drag exerted by the vegetation on the flow. 
We show this by developing a self-consistent model of the flow past rigid vegetation. 
Linear stability analysis of the steady flow resulting from this model naturally leads to the threshold condition, and explains its mechanistic origin. 
Asymptotic analysis of the model for dense grass reveals two unstable modes. 
One of these modes shares many common features with the Kelvin-Helmholtz instability, but is demonstrated to be different from it. 
The predictions of the model for threshold condition and the resulting dominant frequency of waving agree with laboratory observations. 
\end{abstract}
\maketitle
%\section{Introduction}
Sea grasses exhibit a rich set of dynamical behavior due to their collective interaction with both steady and oscillatory flows.  
The resulting changes in hydrodynamic conditions can influence number of environmental processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production \cite{Fonseca87,Grizzle96,Nepf99,Nepf2012}. 
The simplest response of these flexible grass canopies to steady currents is the formation of coherent large amplitude oscillations, known as \monami \cite{AckermanOkubo93}.  
In this letter we provide a hydrodynamic mechanism for the onset of these coherent oscillations.
\newline
Current explanations of \monami invoke the existence of a shear layer near the top of the canopy \cite{Ghisal02,Raupach96} and its instability through a mechanism similar to the Kelvin-Helmholtz instability. 
This instability is thought to lead to coherent eddies over the sea grass canopy which then drives large amplitude synchronous oscillations.
\newline
While the shear layer model appears to predict the frequency of \monami consistent with experimental observations, several aspect of existing theory remain unsatisfactory. 
(i) While the free shear layer is established due to the vegetation drag, linear stability analysis\cite{Raupach96} does not account for this, rendering this explanation inconsistent. 
(ii) Classical free shear flow is known to be unstable for all Reynolds number\cite{drazin}, whereas observations in the lab\cite{Ghisal02} and the field\cite{Grizzle96} indicate the existence of a threshold 
flow speed below which \monami is not observed. 
(iii)The thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the grass tip, is in many cases comparable to the unvegetated layer thickness,
and therefore inconsistent with the assumptions of the free shear layer instability.
Here we present a self-consistent mathematical model that account correctly for these effects 
while also explaining lab experiments and field observations.
%Using this model, the synchronous waving may be understood to be arising from the linear hydrodynamic instability of a uniform flow over the grass bed, such that the principal role of the vegetation is to exert a drag on the flow. 
%A threshold flow speed arises naturally for the onset of the instability; this threshold condition and the resulting waving frequency compares well with experimental observations.
%The vegetation drag introduces a damping mechanism in the flow, in addition to the viscous damping, and overcoming this damping is the primary consideration that sets the threshold flow.
%We deduce this numerically and through an asymptotic scaling analysis in the limit of dense grass.
\newline
The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying \monami because it realizes the shear layer structure of the flow. 
The flexibility of the grass blades, on the other hand, serves merely aid in visualizing the flow structure resulting from the instability\cite{Nepf2012}. 
It has been demonstrated that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels\cite{Ghisal02}. 
Therefore, for simplicity, we consider the grass bed to form a rigid structure that exchanges momentum with the surrounding fluid but does not deform. 
Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated\cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, 
the flow instability is similar to the instability of a shear layer. 
%\begin{figure}
%\begin{center}
%\includegraphics[]{Critical_Re_vs_delta_noshear} \\
%\vspace{-6mm} \hspace{-3mm}
%\includegraphics[]{K_vs_shear_width_noshear}
%\end{center}
%\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density
% parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are 
%also included in the top panel. Inset compares experimental observations 
%of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The %experimental data in the inset in obtained from publications by 
%Ghisalberti and Nepf \cite{Ghisal02} and Vivoni \cite{Vivoni98}. In order to estimate the $\Rey$ for these experiments, a %representative value of $\mu=0.1$ Pa~s was assumed.}
%\label{Re_vs_delta}
%\end{figure}
We use a mean field model for the coupling between the flow and the canopy. 
The force exerted by the grass is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$,  
where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area. In the canopy the equations of continuity \& momentum balance can be written as 
%in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} as
\begin{equation}
\quad \grad\cdot\bu = 0,\hspace{3mm} \rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad P+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
\label{navier-stokes}
\end{equation}
%where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area, 
where $\rho$ the fluid density, $\mathbf{u}$ the velocity, 
$P$ the pressure, $\mu$ the turbulent viscosity and $\mathbf{g}$ the acceleration due to gravity. 
The Reynolds number of the flow based on the scale of a grass blade is expected to be O($10^2$-$10^3$), and hence the drag force is dominated by pressure and modeled as 
$\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn + C_{T}\rho u_{T}^{2}d\bt$ \shreyas{citation needed} 
where $d$ is average width of grass blade $C_{N}$ and $C_{T}$ are the normal and tangential drag coefficients respectively; $u_{T}$, $u_{N}$ are velocity components along and
normal to grass while $\bt,\bn$ being unit vector along and normal to grass. 
We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. 
In the field, both $C_N$ and $N_g$ vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N$ and $N_g$ to be constants. 
Based on previous experiments\cite{Vivoni98}, we take $C_N \approx 1$.
\newline
To understand the onset of \monami, we first calculate the steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ as a solution of \eqref{navier-stokes}, and examine its linear stability. 
Such a steady state flow driven by constant pressure gradient satisfies 
\begin{equation}
 -{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0
\label{base_equ}
\end{equation}
\begin{figure}
\includegraphics[scale=1]{Grass_Base_Nepf_shear}
\caption{Comparison of steady flow profile from the experiments by Ghisalberti and Nepf\cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 
plant height = 13.7$\pm 0.2$ cm and blade width of 0.64 cm)
 and it's approximation with $U_0=7.28$ cm/s and $\delta = 5.02$ cm with our model. The canopy extends upto $y=h_g$ in the water column of depth $2H$. 
The steady velocity profile can be decomposed into three regions, (i) A parabolic profile in the unvegetated region, (ii) a uniform profile deep within the grass bed, and (iii) a boundary layer of thickness $\delta$ near the canopy top. 
The dependence of the boundary layer thickness (estimated numerically as $|U/U_y|$ at $y=\hg$) on the vegetation density parameter $\Rey \Ndg$ is shown in the inset.}
\label{basicflow}
\end{figure}
where $S(y)=1$ in the vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< 2H$. 
In the limit of large grass density, shear stress exerted by the bottom surface is estimated to be negligible in comparison with the vegetation drag and hence the bottom surface ($y=0$) is approximated by zero shear.   
A comparison of laboratory mean profile with the steady flow profile is shown in figure \eqref{basicflow}.
Flow within grass has approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from the balance between drag and with pressure gradient. 
Flow outside has a simple parabolic velocity profile due to the balance between viscous forces and the pressure gradient. 
Continuity of flow speed and shear stresses at canopy top gives rise to a boundary layer near the tip of the grass. 
\newline
This boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and we identify it as a free shear layer\cite{Ghisal02} in the previous explanation of \monami. 
In our model the balance between viscous forces and the drag force in the shear layer 
together with the continuity of shear stress across the grass tip leads to a shear layer thickness $\delta=(\Rey\tilde{N_g})^{-1/3}H$.
% and velocity $u \sim \delta U_0$ in the shear layer. 
Where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, and $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity. 
This dependence of the boundary layer thickness on the grass density $N_g$, shown in Figure \ref{basicflow}, gives us a way to systematically control it and investigate its effect on the instability mechanism.
The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$; minor departures from this scaling can be seen in the inset of figure \ref{basicflow}.
Considering small perturbations $u, v, p$ about the base profile $U$ and $P$ given by the solution of \eqref{base_equ}. Substituting them in equation \eqref{navier-stokes} and expanding  to linear order in the perturbation variable we obtain.
\begin{equation}
\begin{split}
\rho(u_t+U u_x+vU_y) &= -p_x+ {\mu}\nabla^2u-2S\rho C_{N}dN_{g}Uu, \\
\rho(v_t+ Uv_x) &= -p_y+ {\mu}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0.
\end{split}
\end{equation}
These equations can be non-dimensionalized using half channel height $H$, velocity $U_0$, and the associated advection time $H/U_0$. 
We are left with three non-dimensional parameters in the model, ie. $\Rey$, $\Ndg$, and the submergence ratio of the grass $h_g/H$. 
We also use $\delta/H$ in lieu of $\Ndg$ to parametrize the grass density and help elucidate the mechanism of the instability. 
With these scaling, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of 
the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \hat\phi(y) \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation. 
The growth rate $\sigma$ for a given wave number $k=2\pi /\lambda$ can be shown to satisfy the eigenvalue problem.
% provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$ through the solution of
\begin{figure}
\begin{center}
\includegraphics[]{Critical_Re_vs_delta_noshear} \\
\vspace{-6mm} \hspace{-3mm}
\includegraphics[]{K_vs_shear_width_noshear}
\end{center}
\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density
 parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are 
also included in the top panel. Inset compares experimental observations 
of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The experimental data in the inset in obtained from publications by 
Ghisalberti and Nepf \cite{Ghisal02} and Vivoni \cite{Vivoni98}. In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.}
\label{Re_vs_delta}
\end{figure}
\begin{equation}
\begin{split}
\left(D^2 -k^{2} \right)^2\phi &= \Rey \left[ \left({\sigma}+ikU\right) \left(D^2-k^2\right) -ikU_{yy}\right]\phi \\
&+D\left(2R \Ndg S U D \phi\right),
\label{Orr-somerfield}
\end{split}
\end{equation}
where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$.

@@

\begin{table*}
\rowcolors{3}{tableShade}{white}  %% start alternating shades from 3rd row
\renewcommand{\arraystretch}{1.4}
 \begin{tabular}{l|c|c|c}
			& Kelvin-Helmholtz 				& Mode 1 		& Mode 2 \\ \hline
 Base velocity profile 	& $U(y) = U_0 \tanh(y/\delta)$			& \multicolumn{2}{c}{$-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$} \\
 Domain 		& $-\infty < y < \infty$			& \multicolumn{2}{c}{$-1<y<1$} \\
 Inflection point	& exists at $y=0$				& \multicolumn{2}{c}{$U''(y)$ discontinuous at $y=\hg$} \\
 Shear layer thickness	& $\delta$					& \multicolumn{2}{c}{$\delta \sim  H\left(\Rey \Ndg \right)^{-1/3}$} \\
 Linearized dynamics	& Inviscid Rayleigh's equation			& \multicolumn{2}{c}{Equation \eqref{Orr-somerfield}} \\
 Critical parameters	& none						& $\Rey \propto \Ndg^{1/2}$ 	& $\Rey \propto \Ndg$ \\
 Most unstable $k$ as $\delta \to 0$	& $\propto \delta^{-1}$		& $\propto \delta^{-1}$	& $O(1)$ \\
 Mode localized?	& yes, near $y=0$				& ~~~~yes, near $y=\hg$~~~~			& no
 \end{tabular}
 \caption{Comparison between Kelvin-Helmholtz instability and the two modes resulting from solution of \ref{Orr-somerfield}.}
 \label{tab:comparison}
\end{table*}
To better understand the mechanism of waving and the origin of the threshold, we consider the behavior of the instability as a function of the vegetation drag.
The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical grass density discontinuously jumps and remains $O(1)$ (see figure \ref{Re_vs_delta}). 
To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in figure \ref{K_Re_sigma_set3}. 
The lowest $\Rey$ on the neutral curve sets the threshold. 
We observe that as $\Ndg$ increases, the unstable region splits into two. We refer to the unstable region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''. 
The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the grass density increases, causing the discontinuity in the most unstable mode.
